\bf Nonparametric Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test for Marginal Distributions of Stationary Time Series

### 2024-01-15 manuscript/manuscript.tex
comments/edits

\documentclass[12pt, letterpaper]{article}
%\documentclass[12pt, letterpaper, titlepage]{article}

@@

\title{\bf Nonparametric Bootstrap Kolmogorov-Smirnov Test for
  Goodness-of-Fit of Marginal Distributions of Stationary Time Series}
\if0\blind
{
  \author{Mathew Chandy, %\\
%   \href{mailto:mathew.chandy@uconn.edu}
%   {\nolinkurl{mathew.chandy@uconn.edu}}\\
  Elizabeth Schifano\\
  Jun Yan, %\\
  Xianyang Zhang\\
} \fi

@@

@@

Consider a stationary time series $\{X_t: t = 1, \ldots, n\}$ with length~$n$.
Let $H_0: X_t \sim F$, where $F$ is a hypothesized marginal distribution with 
unknown 
parameters. Let $H_A$ be that the marginal distribution of $X_t$ does not follow 
$F$. This is challenging 
because we are
dealing with a situation with both unknown parameters and serial dependence.
If $X_t \sim F$, the true marginal distribution of $X_t$ has some parameters 
$\theta$. The KS goodness-of-fit statistic is to test for the marginal 
distribution of a stationary series.

@@

Let us first consider the case where $X_i$'s are i.i.d. Denote by $F^{(b)}_n$ 
the
empirical distribution of the $b$th bootstrap sample and let
$\hat\theta^{(b)}_n$ be the parameter estimate based on the $b$th bootstrap 
sample. 
Using the bootstrap (asymptotic) theory, we can approximate the distribution of
$\sqrt{n}(F_n(x) - F(x))$ and $\sqrt{n}(F(x; \hat\theta_n) - F(x))$
by that of $\sqrt{n}(F^{(b)}_n(x) - F_n(x))$
and
$\sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; \hat\theta_n))$, respectively.
Therefore, if we define
\begin{equation*}
Y^{(b)}_n(x) = \sqrt{n}(F^{(b)}_n(x) - F_n(x)) - 
\sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; \hat\theta_n)) 
= \sqrt{n}(F^{(b)}_n(x) - F(x; \hat\theta^{(b)}_n)) - 
\sqrt{n}(F_n(x) - F(x; \hat\theta_n)),
\end{equation*}
then $T^{(b)}_n := \sup_x|Y^{(b)}_n(x)|$ is the bootstrap statistic that is 
expected
to approximate the distribution of $T_n$. We note that the term
$\sqrt{n}(F_n(x) - F(x; \hat\theta_n))$ is exactly the bias term considered in 
\citet{babu2004goodness}.

We now consider the case where $X_i$'s are realizations from a time series and
$X^{(b)}_1,...,X^{(b)}_n$ are generated by block bootstrap for 
$b \in \{1, \ldots, B\}$.

Block-bootstrap can be done with overlapping or moving blocks.
Define 
moving blocks (assuming $l > 1$) as:
\begin{equation*}
Z_j =
    \begin{cases}
        \{X_j, \ldots, X_{j + l - 1}\}, & \text{if } j = 1, \dots, n - l + 1,\\
        \{X_j, \ldots, X_n, X_1, \ldots, X_{j-n+l-1}\}, & \text{if } j = n - l
        + 2 ,\dots, n.
    \end{cases}
\end{equation*}

A common 
function for block size that is considered optimal is 
$l = \lceil n^{1/3} \rceil$ \citep{buhlmann1999block},  
which was adopted in this study.

Let $E^{(b)}$ refer to the expectation with respect to
the bootstrap distribution. In this case, we can 
use the distribution of $\sqrt{n}(F^{(b)}_n(x) - E^{(b)}[F^{(b)}_n(x)])$ 
as an approximation of the distribution of
$\sqrt{n}(F_n(x) - F(x))$,
and the distribution of 
$\sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; E^*[\hat\theta^{(b)}_n]))$ to 
approximate $\sqrt{n}(F(x; \theta_n) - F(x))$'s distribution.

We can compute $E^{(b)}[F^{(b)}_n(x)]$ and 
$E^{(b)}[\hat\theta^{(b)}_n]$ numerically (they can also be computed analytically, 
depending on the types of block bootstrap we use). One can compute 
$F^{(b)}_n$ 
and $\hat\theta^{(b)}_n$ based on
the $b$th bootstrap sample. Then
$E^{(b)}[F^{(b)}_n(x)] \approx \frac{1}{B}\sum_{b = 1}^BF^{(b)}_n(x)$, and
$E^{(b)}[\hat\theta^{(b)}_n] \approx \frac{1}{B}\sum_{b = 1}^B\hat\theta^{(b)}_n$.
In this case, we can define
\begin{align*}
  Y^{(b)}_n(x) &= \sqrt{n}(F^{(b)}_n(x) - E^{(b)}[F^{(b)}_n(x)]) - 
             \sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; E^{(b)}[\hat\theta^{(b)}_n)]) \\
           &= \sqrt{n}(F^{(b)}_n(x) - F(x; \hat\theta^{(b)}_n)) -
             \sqrt{n}(E^{(b)}[F^{(b)}_n(x)] - F(x; E^{(b)}[\hat\theta^{(b)}_n])),
\end{align*}
and $T^{(b)}_n = \sup_x|Y^{(b)}_n(x)|$.

@@

\paragraph{Size}
We first must evaluate whether this method works when $H_0$ is true. To
test this, we can
generate a simulated sample $X_t$ with a certain marginal distribution 
$F_\theta$,
and use our method to test if $X_t$ indeed has the marginal distribution $F$ 
with some unknown $\theta$. If the test holds its size, the 
p-value
of the test should be uniformally distributed between 0 and 1. We must try the
method with different marginal distributions to ensure that it is robust.
In order for the method to work, a large sample size may be necessary.

We generated time series with marginal distributions $N(8, 8)$ and 
$\Gamma(8, 1)$, lag-1 autocorrelation $\phi \in \{-.4, -.2, 0, .2, .4\}$, and
sample size $n \in \{100, 200, 400, 600, 800, 1600, 3200\}$. For the purposes of 
evaluating if the test holds it size, when $X_t \sim N(8, 8)$, we tested that the 
marginal distribution family is normal, and when $X_t \sim \Gamma(8, 1)$, we tested
that the marginal distribution family is gamma.
For each setting for $F$, $\phi$, and $n$, we replicate the method 1000 times 
to get the distribution of the p-values 
for the test when applied to samples from the same data generating process.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/normal}
  \caption{A Q-Q plot of the p-values testing that a distribution
  generated from a $N(8,8)$ data generating process is normal.}
  \label{fig:normal}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[scale=1]{figures/zoom_normal}
  \caption{A Q-Q plot of the p-values displayed in Figure~\ref{fig:normal} zoomed in to 
  probabilities between 0 and
  0.1}
  \label{fig:zoom_normal}
\end{figure}

@@

\paragraph{Power}
We must also evaluate if this method works when $H_0$ is false. We can use
the block bootstrap KS test to test if some $X_t \sim F$ is generated from 
some other 
distribution $G$ with some unknown $\theta$. In this scenario, if the test is 
indeed powerful,
we would ideally want $\beta$ 
(or the probability of failed rejection under $H_A$) 
to be 0, but we
expect it to be generally low. Again, we must try the method with different
marginal distributions to ensure that it is robust.
We would then expect the distribution of p-values to be non-uniform, and the 
rate
of $\#P < \alpha$ to be high, meaning a higher density of low p-values.

@@

@@

@@

We expect the marginal distribution of Microsoft stock returns to have a 
slightly
heavier tail than the normal distribution. Using the \textsl{tseries} package 
\citep{tseries}, 
we gathered closing stock prices from January 1st, 1998 to December 31st, 2022.
Then we approximated stock returns by taking the difference in the logarithm 
of the closing prices.
We tested that the stock returns are normally
distributed (which we expect to reject).  After that, we tested that the
marginal distribution of the time series follows the Student's $t$ distribution.
While we 
could choose to estimate a fitted value of the degrees of freedom parameter $v$
as
part of our method, it is very difficult to do so. Therefore, we attempted to
test 
if it follows
the location-scale $t$ distribution with increasingly smaller specified degrees 
of freedoms. This was achieved with the \textsl{extraDistr} 
package \citep{extraDistr}. We
also observed how the method performs using increasingly smaller durations of
time. First, we tred the entire data set which has a duration of 25 years. Then,
we used the last 5 years of data: January 1st, 2018 to December 31st, 2022.
Finally, we tried the method only using the last year of data: January 1st, 2022
to December 31st, 2022.

@@

@@

>>>>

\documentclass[12pt, titlepage, letterpaper]{article}
%\documentclass[12pt, letterpaper, titlepage]{article}

@@

\title{\bf Nonparametric Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test for
  Marginal Distributions of Stationary Time Series}
\if0\blind
{
  \author{Mathew Chandy, %\\
%   \href{mailto:mathew.chandy@uconn.edu}
%   {\nolinkurl{mathew.chandy@uconn.edu}}\\
  Elizabeth Schifano\\
  Jun Yan, %\\
  Xianyang Zhang\\
} \fi

@@

\jy{Start from the need for KS test for stationary series; review existing works
  and identify the gap; summarize the contribution.}

@@

Consider a stationary time series $\{X_t: t = 1, \ldots, n\}$ with length~$n$.
We are interested in testing whether or not $X_t$ follows a distribution in a
parametric family of distribution~$F$ indexed by a parameter
vector~$\theta$. That is, the null hypothesis is
\[
  H_0: X_t \sim F(\cdot \mid \theta), \quad t = 1, \ldots, n,
\]
for some unspecified parameter $\theta$.
The alternative hypothesis $H_A$ is that the marginal distribution of $X_t$ does
not follow~$F$ for any parameter value~$\theta$. This is a challenging situation
both the parameters and the serial dependence structure are unknown.

\jy{Need a better opening sentence. Notation $\hat\theta_n$ and $F_n$ have
  not been defined.}

@@

Let us first consider the case where $X_i$'s are i.i.d. Denote by $F^{(b)}_n$ 
the empirical distribution of the $b$th bootstrap sample and let
$\hat\theta^{(b)}_n$ be the parameter estimate based on the $b$th bootstrap 
sample. 
Using the bootstrap (asymptotic) theory, we can approximate the distribution of
\jy{need references to back up}
$\sqrt{n}(F_n(x) - F(x))$ and $\sqrt{n}(F(x; \hat\theta_n) - F(x))$
by that of $\sqrt{n}(F^{(b)}_n(x) - F_n(x))$
and
$\sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; \hat\theta_n))$, respectively.
Therefore, if we define
\begin{align*}
Y^{(b)}_n(x) &= \sqrt{n}(F^{(b)}_n(x) - F_n(x)) - 
               \sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; \hat\theta_n)) \\
             &= \sqrt{n}(F^{(b)}_n(x) - F(x; \hat\theta^{(b)}_n)) - 
               \sqrt{n}(F_n(x) - F(x; \hat\theta_n)),
\end{align*}
then $T^{(b)}_n := \sup_x|Y^{(b)}_n(x)|$ is the bootstrap statistic that is 
expected
to approximate the distribution of $T_n$. We note that the term
$\sqrt{n}(F_n(x) - F(x; \hat\theta_n))$ is exactly the bias term considered in 
\citet{babu2004goodness}.

We now consider the case where $X_i$'s are realizations from a time series and
$X^{(b)}_1,...,X^{(b)}_n$ are generated by block bootstrap for 
$b \in \{1, \ldots, B\}$.  
Block-bootstrap can be done with overlapping or moving blocks.
Define 
moving blocks (assuming $l > 1$) as:
\begin{equation*}
Z_j =
    \begin{cases}
        \{X_j, \ldots, X_{j + l - 1}\}, & \text{if } j = 1, \dots, n - l + 1,\\
        \{X_j, \ldots, X_n, X_1, \ldots, X_{j-n+l-1}\}, & \text{if } j = n - l
        + 2 ,\dots, n.
    \end{cases}
\end{equation*}
A common 
function for block size that is considered optimal is 
$l = \lceil n^{1/3} \rceil$ \citep{buhlmann1999block},  
which was adopted in this study.

\jy{Not good opening sentence}
Let $E^{(b)}$ refer to the expectation with respect to
the bootstrap distribution. In this case, we can 
use the distribution of $\sqrt{n}(F^{(b)}_n(x) - E^{(b)}[F^{(b)}_n(x)])$
\jy{The second term: does it depend on $b$? }
as an approximation of the distribution of
$\sqrt{n}(F_n(x) - F(x))$,
and the distribution of 
$\sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; E^*[\hat\theta^{(b)}_n]))$ to
approximate $\sqrt{n}(F(x; \theta_n) - F(x))$'s distribution.

We can compute $E^{(b)}[F^{(b)}_n(x)]$ and
\jy{Are these two terms $b$-specific?}
$E^{(b)}[\hat\theta^{(b)}_n]$ numerically (they can also be computed analytically, 
depending on the types of block bootstrap we use). One can compute 
$F^{(b)}_n$ 
and $\hat\theta^{(b)}_n$ based on
the $b$th bootstrap sample. Then
$E^{(b)}[F^{(b)}_n(x)] \approx \frac{1}{B}\sum_{b = 1}^BF^{(b)}_n(x)$, and
$E^{(b)}[\hat\theta^{(b)}_n] \approx \frac{1}{B}\sum_{b = 1}^B\hat\theta^{(b)}_n$.
In this case, we can define
\begin{align*}
  Y^{(b)}_n(x) &= \sqrt{n}(F^{(b)}_n(x) - E^{(b)}[F^{(b)}_n(x)]) - 
             \sqrt{n}(F(x; \hat\theta^{(b)}_n) - F(x; E^{(b)}[\hat\theta^{(b)}_n)]) \\
           &= \sqrt{n}(F^{(b)}_n(x) - F(x; \hat\theta^{(b)}_n)) -
             \sqrt{n}(E^{(b)}[F^{(b)}_n(x)] - F(x; E^{(b)}[\hat\theta^{(b)}_n])),
\end{align*}
and $T^{(b)}_n = \sup_x|Y^{(b)}_n(x)|$.

@@

\subsection{Size}
We first must evaluate whether this method works when $H_0$ is true. To
test this, we can
generate a simulated sample $X_t$ with a certain marginal distribution 
$F_\theta$,
and use our method to test if $X_t$ indeed has the marginal distribution $F$ 
with some unknown $\theta$. If the test holds its size, the 
p-value
of the test should be uniformally distributed between 0 and 1. We must try the
method with different marginal distributions to ensure that it is robust.
In order for the method to work, a large sample size may be necessary.

\jy{Give details on how the data were generated. Are these enough infomation for
  people reproduce the results? I'd suggest that we use Kendall's tau to measure
  the serial dependence. }

\jy{How many replicates were used to estimate the size and power?}

We generated time series with marginal distributions $N(8, 8)$ and
\jy{why choose these two distributions?}
$\Gamma(8, 1)$, lag-1 autocorrelation $\phi \in \{-.4, -.2, 0, .2, .4\}$, and
sample size $n \in \{100, 200, 400, 600, 800, 1600, 3200\}$. For the purposes of 
evaluating if the test holds it size, when $X_t \sim N(8, 8)$, we tested that the 
marginal distribution family is normal, and when $X_t \sim \Gamma(8, 1)$, we tested
that the marginal distribution family is gamma.
For each setting for $F$, $\phi$, and $n$, we replicate the method 1000 times 
to get the distribution of the p-values 
for the test when applied to samples from the same data generating process.

@@

\jy{Reduce width of  the panel title section. Reduce the font size of the
  labels. Generate the figures with the right size in mind. Use whole textwidth.}

\begin{figure}[tbp]
  \centering
  \includegraphics[width = \textwidth]{figures/normal}
  \caption{A Q-Q plot of the p-values testing that a distribution
  generated from a $N(8,8)$ data generating process is normal.}
  \label{fig:normal}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width = \textwidth]{figures/zoom_normal}
  \caption{A Q-Q plot of the p-values displayed in Figure~\ref{fig:normal} zoomed in to 
  probabilities between 0 and
  0.1}
  \label{fig:zoom_normal}
\end{figure}

@@

\subsection{Power}
We must also evaluate if this method works when $H_0$ is false. We can use
the block bootstrap KS test to test if some $X_t \sim F$ is generated from 
some other 
distribution $G$ with some unknown $\theta$. In this scenario, if the test is 
indeed powerful,
we would ideally want $\beta$ 
(or the probability of failed rejection under $H_A$) 
to be 0, but we
expect it to be generally low. Again, we must try the method with different
marginal distributions to ensure that it is robust.
We would then expect the distribution of p-values to be non-uniform, and the 
rate
of $\#P < \alpha$ to be high, meaning a higher density of low p-values.

@@

\jy{Clean up the tables. Use percentage for power. Keep three effective
  digits. Why are there so many decimal places? }

\jy{Only three sample sizes will be enough. Sample size 800 or higher give
  power~1. Only report nominal size 0.05 would be ok. May consider presenting
  the intervals in a figure.}

@@

\jy{No need for such small a table. Could plot the histogram and fitted density.}

@@

We expect the marginal distribution of Microsoft stock returns to have a 
slightly
heavier tail than the normal distribution. Using the \textsl{tseries} package 
\citep{tseries}, 
we gathered closing stock prices from January 1st, 1998 to December 31st, 2022.
Then we approximated stock returns by taking the difference in the logarithm 
of the closing prices.
We tested that the stock returns are normally
distributed (which we expect to reject).  After that, we tested that the
marginal distribution of the time series follows the Student's $t$ distribution
with a fixed degrees of freedom.
While we 
could choose to estimate a fitted value of the degrees of freedom parameter $v$
as
part of our method, it is very difficult to do so. Therefore, we attempted to
test 
if it follows
the location-scale $t$ distribution with increasingly smaller specified degrees 
of freedoms. This was achieved with the \textsl{extraDistr} 
package \citep{extraDistr}. We
also observed how the method performs using increasingly smaller durations of
time. First, we tred the entire data set which has a duration of 25 years. Then,
we used the last 5 years of data: January 1st, 2018 to December 31st, 2022.
Finally, we tried the method only using the last year of data: January 1st, 2022
to December 31st, 2022.

@@

\jy{Normal was not rejected for 1 year. How about 2 year, 3 year, 4 year?}

@@

\jy{Quality control the bib source}
